---
title: "Drug Group Analyzer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Libraries

```{r}
library(tidyverse)
```

Load Diabetes Data from the Excel

```{r}
file_path = "/Users/prolap/v2/hult/events/HultHustle/HultHustle/files/diabetic_data.xlsx"
```

Read Diabetes Data

```{r}
diabetic_data_df = read_excel(file_path, sheet=1)
# Clean Diabetes Main Data
diabetic_data_df =diabetic_data_df %>% mutate(age = str_replace_all(age, "\\[|\\)", ""))
diabetic_data_df[diabetic_data_df == "?"] <- NA
glimpse(diabetic_data_df)
```

Read Drug Mapper Data

```{r}
drug_type_mapper_df = read_excel(file_path, sheet=2, col_names=c('drug', 'drug_group'))
glimpse(drug_type_mapper_df)
```


```{r}
drugs_df = diabetic_data_df %>% select(admission_type_id, admission_source_id, medical_specialty, (25:50))
drugs_df
```

```{r}
drug_types = c("Biguanides", "Sulfonylureas", "Alpha-glucosidase", "Meglitinides", "Thiazolidinediones", "gliptins")
for (drugtype in drug_types) {
  print(drugtype)
  v_drug_list = drug_type_mapper_df %>% filter(drug_group == drugtype) %>% select("drug")  %>% unlist(use.names = FALSE)
  print(length(v_drug_list))
}

```
	
	Create Columns for each of the Drug Groups so that we can run analytics over them.

```{r}
v_drug_list1 = drug_type_mapper_df %>% filter(drug_group == "Biguanides") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(Biguanides = ifelse(.data[[v_drug_list1[[1]]]] != "No" |
                                        .data[[v_drug_list1[[2]]]] != "No" |
                                        .data[[v_drug_list1[[3]]]] != "No", 
                                        "Yes", "No"))

v_drug_list2 = drug_type_mapper_df %>% filter(drug_group == "Sulfonylureas") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(Sulfonylureas = ifelse(.data[[v_drug_list2[[1]]]] != "No" |
                                        .data[[v_drug_list2[[2]]]] != "No" |
                                        .data[[v_drug_list2[[3]]]] != "No" |
                                        .data[[v_drug_list2[[4]]]] != "No" |
                                        .data[[v_drug_list2[[5]]]] != "No" |
                                        .data[[v_drug_list2[[6]]]] != "No" |
                                        .data[[v_drug_list2[[7]]]] != "No" |
                                        .data[[v_drug_list2[[8]]]] != "No" |
                                        .data[[v_drug_list2[[9]]]] != "No", 
                                        "Yes", "No"))

v_drug_list3 = drug_type_mapper_df %>% filter(drug_group == "Alpha-glucosidase") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(AlphaGlucosidase = ifelse(.data[[v_drug_list3[[1]]]] != "No" |
                                        .data[[v_drug_list3[[2]]]] != "No",
                                        "Yes", "No"))


v_drug_list4 = drug_type_mapper_df %>% filter(drug_group == "Meglitinides") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(Meglitinides = ifelse(.data[[v_drug_list4[[1]]]] != "No" |
                                        .data[[v_drug_list4[[2]]]] != "No",
                                        "Yes", "No"))

v_drug_list5 = drug_type_mapper_df %>% filter(drug_group == "Thiazolidinediones") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(Thiazolidinediones = ifelse(.data[[v_drug_list5[[1]]]] != "No" |
                                        .data[[v_drug_list5[[2]]]] != "No", 
                                        "Yes", "No"))

v_drug_list6 = drug_type_mapper_df %>% filter(drug_group == "gliptins") %>% select("drug")  %>% unlist(use.names = FALSE)
drugs_df = drugs_df %>% mutate(gliptins = ifelse(.data[[v_drug_list[[1]]]] != "No", 
                                        "Yes", "No"))

```

Quick Clean up over medicines

```{r}
drugs_df <- drugs_df[-c(4:27)]
drugs_df
```

```{r}
col_order <- c("Biguanides", "Sulfonylureas", "AlphaGlucosidase", "Meglitinides", "Thiazolidinediones", "gliptins", "admission_type_id", "admission_source_id", "medical_specialty", "readmitted")
drugs_df <- drugs_df[, col_order]
drugs_df
```

Group Plot

```{r}
drug_groups <- c("Biguanides", "Sulfonylureas", "AlphaGlucosidase", "Meglitinides", "Thiazolidinediones", "gliptins")
plot_df = as.data.frame(table(drugs_df[drug_groups[1]]))
ggplot(plot_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title=drug_groups[1], x = "", y = "Frequency")
plot_df = as.data.frame(table(drugs_df[drug_groups[2]]))
ggplot(plot_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title=drug_groups[2], x = "", y = "Frequency")
plot_df = as.data.frame(table(drugs_df[drug_groups[3]]))
ggplot(plot_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title=drug_groups[3], x = "", y = "Frequency")
plot_df = as.data.frame(table(drugs_df[drug_groups[4]]))
ggplot(plot_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title=drug_groups[4], x = "", y = "Frequency")
plot_df = as.data.frame(table(drugs_df[drug_groups[5]]))
ggplot(plot_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title=drug_groups[5], x = "", y = "Frequency")
```
From the data set it is evident that Sulfonylureas are administered the most followed by Biguanides. Lets see the efficiency of these 
drug groups

Efficacy of Sulfonylureas

```{r}
drugs_df <- drugs_df %>% mutate(readmitted = as.factor(readmitted), Sulfonylureas = as.factor(Sulfonylureas), Biguanides=as.factor(Biguanides))
# Testing the efficacy of Sulfonylureas
plot_df = drugs_df %>% filter(Sulfonylureas == "Yes")
graph1_df = as.data.frame(table(plot_df$readmitted))
graph1_df
ggplot(graph1_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title="Sulfonylureas - Readmission", x = "", y = "Frequency")
```
Efficacy of Sulfonylureas

```{r}
# Custom Parameters <30 
efficacy_sulfonylureas = 15276 / (15276+3159+10510)
print(efficacy_sulfonylureas*100)
```


Efficacy of Biguanides

```{r}
# Testing the efficacy of Biguanides
plot_df = drugs_df %>% filter(Biguanides == "Yes")
graph2_df = as.data.frame(table(plot_df$readmitted))
graph2_df
ggplot(graph2_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + labs(title="Biguanides - Readmission", x = "", y = "Frequency")
```

```{r}
efficacy_biguanides = 11344 / (11344+6706+1939)
print(efficacy_biguanides*100)
```


In our research efficacy of Biguanides for hospitalized case with respect to this dataset with readmission factor came up to 56.75% whereas efficacy of Sulfonylureas was 52.78%. This would be a great point to start while making RND decisions for a new drug for Type 2 Diabetes


























