---
title: "Data Analytics over Diabetes Admission Data"
output: html_notebook
---

Import Libraries

```{r}
library(tidyverse)
library(readxl)
library("dplyr")
library("ggplot2")
library(stringr)
library(assertive)
```

Load Diabetes Data from the Excel

```{r}
DIABETES_DATA = "/Users/prolap/v2/hult/events/HultHustle/HultHustle/files/diabetic_data.xlsx"
```


```{r}
diabetic_data = read_excel(DIABETES_DATA, sheet=1)
```

Clean Diabetes Data - Remove square brackets and modify the question mark to ?

```{r}
# Clean Diabetes Main Data
diabetic_data =diabetic_data %>% mutate(age = str_replace_all(age, "\\[|\\)", ""))
diabetic_data[diabetic_data == "?"] <- NA
glimpse(diabetic_data)
```

Merge Admission Source Id with the diabetes data

```{r}
admission_types_mapper_df = read_excel(DIABETES_DATA, sheet=3)
glimpse(admission_types_mapper_df)
```

Merge - Clean Admission Data and run visualization with the diabetes data

```{r}
# Query Admission Data by removing bad data
admissions_data <- diabetic_data %>% filter(!admission_type_id %in% c(5,6,8))
temp_df <- (table(admissions_data$admission_type_id))
temp_df <- as.data.frame(temp_df)
temp_df <- merge(temp_df, admission_types_mapper_df, by.x ="Var1", by.y = "admission_type_id")
temp_df <- temp_df %>% mutate(Var1 = as.factor(Var1), description = as.factor(description))
assert_is_data.frame(temp_df)

# Plot [Admission Type vs Frequency in the Dataset]
admissions_type_frequency_plot = ggplot(temp_df, aes(x=description, y=Freq)) + 
  geom_bar(stat="identity", fill="#0045a5") + 
  geom_text(aes(label=Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Admission Type") + ylab("Frequency") + scale_y_log10()
admissions_type_frequency_plot
```

Merge Admission Sources with the Diabetes Data and run plots

```{r}
admission_sources_mapper_df = read_excel(DIABETES_DATA, sheet=5)
glimpse(admission_sources_mapper_df)

# Admission Sources Create Summary Table
admission_sources_data <- diabetic_data %>% filter(!admission_source_id %in% c(9,15,17,20,21))
admission_sources_clean_data <- admission_sources_mapper_df %>% filter(!admission_source_id %in% c(9,15,17,20,21))
temp_df <- table(admission_sources_data$admission_source_id)
temp_df <- as.data.frame(temp_df)
assert_is_data.frame(temp_df)
temp_df <- merge(temp_df, admission_sources_clean_data, by.x ="Var1", by.y = "admission_source_id")
temp_df <- temp_df[order (-temp_df[,2]),]  %>% mutate(Freq = as.numeric(Freq))
# Reset Index
rownames(temp_df) = seq(length=nrow(temp_df))
top_admission_sources = temp_df[1:6,] %>% mutate(Var1 = as.factor(Var1), description = as.factor(description))
top_admission_sources

admissions_sources_frequency_plot = ggplot(data=top_admission_sources, aes(x=description, y=Freq)) + 
  geom_bar(stat="identity", fill="#0045a5") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_text(aes(label=Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Admission Source") + ylab("Frequency")
admissions_sources_frequency_plot
```

Factorize Age and Gender

```{r}
# No of the Columns
sprintf("No of Columns %d", length(names(diabetic_data)))
# Clean Age
diabetic_data <- diabetic_data %>% 
    mutate(age = str_replace_all(age, "\\[|\\)", ""),
           age = as.factor(age),
           gender = as.factor(gender))
# Assert gender is a factor
assert_is_factor(diabetic_data$gender)

```


Gender Analysis

```{r}
# Gender
table(diabetic_data$gender)
# There are 3 records with gender with data "Unknown/Invalid" - Let's Remove it 
diabetic_data<-diabetic_data[!(diabetic_data$gender=="Unknown/Invalid"),]
age_gender_plot = ggplot(diabetic_data, aes(x=age, fill=gender)) + geom_bar(position = "dodge") + 
  ggtitle("Age Vs Gender") + xlab("Age") + ylab("Gender")
age_gender_plot
```

Medical Specialty

```{r}
# Medical Specialty
table(diabetic_data$medical_specialty)
```


```{r}
graph_df = as.data.frame(table(diabetic_data$medical_specialty))
graph2_df = graph_df %>% filter(Var1 != "?") %>% arrange(desc(Freq)) %>% slice(1:10)
ggplot(graph2_df, aes(x=Var1, y=Freq, fill=Var1)) + geom_col(width=0.2) + coord_flip() + labs(title="Medical Speciality treating Diabetes", x = "", y = "Frequency")
```

Internal Medicine treated most number of cases followed by Family / General Practice and Emergency/ Trauma

